---
title: "433 hw1"
author: "Bo Li"
date: "2022/9/26"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(nycflights13)
library(ggplot2)
```

## Q1
*How many flights have a missing dep_time? What other variables are missing? What might these rows represent?*
```{r}
filter(flights,is.na(dep_time))
```

With 8255 missing dep_time in the flight data frame.

dep_delay, arr_time, arr_delay, and air_time are also missing.

The flight may be canceled so those variables are NA.

## Q2
*Currently dep_time and sched_dep_time are convenient to look at, but hard to compute with because they’re not really continuous numbers. Convert them to a more convenient representation of number of minutes since midnight.*

```{r}
# Original format = HHMM, so integer division by 100 to get hours and mod by 100 to get minutes.
# Mod by (24*60) to treat start as midnight
flights<- mutate(flights,
  dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% (24*60),
  sched_dep_time = (sched_dep_time %/% 100 * 60 + sched_dep_time %% 100) %% (24*60)
)
flights
```

## Q3
*Look at the number of canceled flights per day. Is there a pattern? Is the proportion of canceled flights related to the average delay? Use multiple dyplr operations, all on one line, concluding with ggplot(aes(x= ,y=)) + geom_point()*

```{r}
flights %>%
  group_by(year,month,day) %>%
  summarise(canceled_per_day= sum(is.na(air_time)),
    mean_arr_delay = mean(arr_delay, na.rm = TRUE),
    mean_dep_delay = mean(dep_delay, na.rm = TRUE)) %>%
  ggplot(aes(y=canceled_per_day)) + 
  geom_point(aes(x=mean_arr_delay)) + 
  geom_point(aes(x=mean_dep_delay))+
  labs(x = "Average delay", y = "Canceled flights")
```

There is a positive correlation between canceled flights and the average delay. As increasing in average delays, the canceled flights also increase.